---
title: "Avocado Data Statistics"
author: "Deondre Brown"
date: "2024-03-03"
output:
  github_document: default
  pdf_document: default
---

This is a project documenting my statistical steps and analysis regarding a dataset for Avocado sales.

Step 1.Read in the data

```{r}
avocado <- read.csv("Module2_Avocado_Data.csv")
```

Step 2. Load in the necessary packages we need
```{r}
library(tidyverse)
library(psych)
```

Step 3. Use these functions to make the sure data was loaded in properly.
```{r}
dim(avocado) # Shows the numbers of rows and columns in the dataset
```

```{r}
names(avocado) # Names all the variables in the dataset
```

```{r}
view(avocado) # Displays the dataset in a spreadsheet format in another tab
```

```{r}
head(avocado, 10) # Prints the first 10 rows of the data
```

```{r}
str(avocado) # Get an overview of the variables and their first rows of values
```

Step 4. Create a new data column to hold dollar sales information. To get the dollar sales information we are going to multiple the average price by the total volume.

```{r}
avocado$totalsales <- avocado$averageprice*avocado$totalvolume
```

Step 5. Look at basic summaries of the data.
```{r}
summary(avocado) #R built-in function for summary statistics
```

```{r}
describe(avocado) #Better than summary() function and gives more statistics
```

```{r}
avocado |>
  select(totalvolume, totalsales) |>
  describe() # Describes statistics for the total volume and total sales variables
```

**Using these statistics we can find out the minimum, maximum, and average sales**

Minimum: **563.35**
Maximum: **40,361,458**
and Average: **665,447.3**

Step 6. Look at a data distribution using a histogram. For this histogram we can display the frequency regarding the average price of avocados.

```{r}
hist(avocado$averageprice, 
     col = "red",
     xlab = "Average Price",
     ylab = "Frequency")
```

**Using this histogram, we can make the observation that the data is positively skewed**

Step 7. We can find the correlation value between average price and sales volume to see how they relate to each other.
```{r}
cor(avocado$averageprice,avocado$totalvolume)
```

**We get a value of -0.31 indicating that the average price and sales volume are inversely related**

Step 8. We can generate many more insights by gathering summary statistics by grouping.

**Get the Summary statistics for price by each supplier**
```{r}
describeBy(averageprice ~ supplierid, mat=T, data=avocado) #Mat function transposes the data
```

**Average price of avocados by month?**
```{r}
describeBy(averageprice ~ month, mat=T, data=avocado)
```


